\magnification=1000%\magstephalf
\parskip3pt
\baselineskip12pt

\centerline{\bf IWLS 2018 programming contest}
\centerline{Mathias Soeken, \sl EPFL \qquad \rm Alan Mishchenko, \sl UC Berkeley}

\bigskip

\noindent {\bf Boolean logic networks.} \enspace We consider single
output Boolean logic networks that realize Boolean functions $f(x_1,
\dots, x_n)$.  A $k$-feasible $n$-input Boolean logic network is a
sequence of gates
%
$$
  x_i = g_i(x_{s(i,1)}, \dots, x_{s(i,k)}) \qquad\hbox{for $n < i \le n+r$,}
$$
where $g_i$ is any $k$-input Boolean function with inputs from
previous gates or primary inputs $x_{s(i,k)}$, with $1 \le s(i,j) < i$
for all $1 \le j \le k$.  The Boolean function realized by the network
is obtained from the value to $x_{n+r}$ for all assignments to the
primary inputs $x_1, \dots, x_n$.

For convenience, we define $S_i = \{s(i,1), \dots, s(i,k)\}$ as the
support variables of gate $i$.

A Boolean logic network is {\it normalized\/}, if all of the following
conditions hold.
\item{(i)} $g_i(0, 0, \dots, 0) = 0$ for all $n < i \le n+r$.
\item{(ii)} $s(i, 1) \le s(i, 2) \le \cdots \le s(i, k)$ for all $n < i \le n+r$.
\item{(iii)} If $s(i+1,k) \neq i$, then $(s(i,1), s(i,2), \dots,
s(i,k)) \preceq (s(i+1,1), s(i+1,2), \dots, s(i+1,k))$.
\item{(iv)} If $S_i = S_{i+1}$, then $g_i \prec g_{i+1}$
\item{(v)} If the primary inputs $x_j$ and $x_l$ are symmetric, then
$\arg\min_{i} \{j \in S_i\} \le \arg\min_i \{l \in S_i\}$.

Note that normalized Boolean logic networks cannot represent all
functions $f$, but only those for which $f(0, 0, \dots, 0) = 0$.

Condition (iii) states, that if two successive gates are independent
from each other (i.e., $s(i+1) \neq i$), then we require the two gates
to be lexicographically ordered based on their support.  For example,
$(1, 3, 3) \prec (2, 2, 3)$, and $(1, 2, 2) \prec (1, 2, 3)$, but $(2,
3, 3) \not\prec (1, 3, 3)$.

Condition (iv) states, that if two successive gates have the same
support, the gates must be ordered according to their functions.  We
do this by considering their truth tables as bitstrings of size $2^k$.
For example, 2-input AND ($1000_2 = 8_{10}$) is smaller than 2-input
OR ($1110_2 = 14_{10}$).

Finally, condition (v) states that if two primary inputs $x_j$ and
$x_l$ are symmetric (i.e., we can swap the two inputs without changing
the function), then input $x_j$ must be used by a gate before input
$x_l$ is used.

\smallskip\noindent{\bf Task.} \enspace Write a computer program, that
for a given a Boolean function $f(x_1, \dots, x_n)$, a number $k$, and
a number of gates $r$, finds as many $k$-feasible normalized Boolean
logic networks as possible that realize $f$ with $r$ gates.  The
program should terminate within one hour and should write all found
logic networks into a single file called {\tt <func>-$k$-$r$.bln}.  In
that filename, {\tt <func>} is the hexadecimal truth table
representation of the function (from the given list of contest
functions).  Each network is described with a line for each gate
formatted as follows:

\medskip
\bgroup
\parindent=40pt
\tt\obeylines
${\rm chr}(i)$ = <binary truth table of $g_i$> ${\rm chr}(s(i,1))$ $\dots$ ${\rm chr}(s(i,k))$
\egroup

\medskip\noindent where ${\rm chr}(i)$ is the letter in the alphabet
at position $i$, i.e., ${\rm chr}(1) = {\tt a}$, ${\rm chr}(2) = {\tt
b}$, and so on.  For convenience, we use lower case letters for
primary inputs ($i \le n$), and upper case letters for gates ($i >
n$).  As an example, the file {\tt e8-2-4.bln} for logic networks of the majority function may contain the 4-gate solution
%
$$
x_4 = x_1 \oplus x_2,\;\; x_5 = x_1 \lor x_2,\;\; x_6 = x_3 < x_4,\;\; x_7 = x_5 > x_6
$$
%
formatted as

\medskip
\bgroup
\parindent=40pt
\parskip=0pt
\baselineskip=12pt
\tt\obeylines
D = 0110 a b
E = 1110 a b
F = 0100 c D
G = 0010 E F
\egroup

\smallskip\noindent All gates should be in order, and multiple
solutions should be seperated from each other by an empty line.  Also
note that $x_{s(i,1)}$ is the least significant input to function
$g_i$.  To illustrate that, the truth table for implication is {\tt
1101}.

\bye
